A Bias-Variance Decomposition of Information Transfer in Multi-Task Learning

### 2020-08-20 arxiv/intro.tex
wip

\begin{abstract}
	Multi-task learning is a powerful approach in many applications such as image and text classification. Yet, there is little rigorous understanding of when multi-task learning outperforms single-task learning. In this work, we provide a rigorous study to anwer the question in the high-dimensional linear regression setting. We show that a bias-variance tradeoff of multi-task learning determines the effect of information transfer and develop new concentration bounds to analyze the tradeoff. Our key observation is that three properties of task data, namely \textit{task similarity}, \textit{sample size}, and \textit{covariate shift} can affect transfer in the high-dimensional linear regression setting. We relate each property to the bias and variance of multi-task learning and explain three negative effects with decreased task similarity,	increased source sample size, and covariate shift under increased source sample size. We validate the three effects on text classification tasks. Inspired by our theory, we show two practical connections of interest.
	First, single-task results can help understand when multi-task learning gives gains. Second, incrementally adding training data can mitigate negative transfer and improve multi-task training efficiency.
\end{abstract}

@@

%\textbf{Setup and Main Results.}
In this work, we study the bias and variance of multi-task learning in the high-dimensional linear regression setting \cite{HMRT19,BLLT20}.
Our key observation is that three properties of task data, including \textit{task similarity}, \textit{sample size}, and \textit{covariate shift}, can affect whether multi-task learning outperforms single-task learning (which we refer to as \textit{positive transfer}).
As an example, we vary each property in Figure \ref{fig_model_shift_phasetrans} for two linear regression tasks and measure the improvement of multi-task learning over single-task learning for a particular task.
We observe that the effect of transfer can be positive or negative as we vary each property.
These phenomena cannot be explained using previous techniques \cite{WZR20}.
The high-dimensional linear regression setting allows us to measure the three properties precisely.
We define each property for the case of two tasks and our definition applies to general settings.
We refer to the first task as the source task and the second as the target task.
\squishlist
	\item \textbf{Task similarity:} Assume that both tasks follow a linear model with parameters $\beta_1, \beta_2\in\real^p$, respectively.
	We measure the distance between them by $\norm{\beta_1 - \beta_2}$.
	\item \textbf{Sample size:} Let $n_1 = \rho_1 \cdot p, n_2 = \rho_2 \cdot p$ be the sample size of each task, where $\rho_1, \rho_2>1$ are both fixed values that do not grow with $p$.
	We measure the source/target sample ratio by $\rho_1 / \rho_2$.
%	Importantly, $\rho_2$ can be a small constant (say $2$) to capture the need for more labeled data.
	\item \textbf{Covariate shift:} Assume that the task features are random vectors with positive semidefinite covariance matrix $\Sigma_1,\Sigma_2\in\real^{p\times p}$, respectively.
	%$x = \Sigma_i^{1/2}z$, where $z\in\real^p$ consists of i.i.d. entries with mean zero and unit variance, and is a positive semidefinite matrix.
	We measure covariate shift with matrix $\Sigma_1^{1/2}\Sigma_2^{-1/2}$.
\squishend

>>>>

\begin{abstract}
	Multi-task learning is a powerful approach in many applications such as image and text classification. Yet, there is little rigorous understanding of when multi-task learning outperforms single-task learning. In this work, we provide a rigorous study to anwer the question in the high-dimensional linear regression setting. We show that a bias-variance tradeoff of multi-task learning determines the effect of information transfer and develop new concentration bounds to analyze the tradeoff. Our key observation is that three properties of task data, namely \textit{task similarity}, \textit{sample ratio}, and \textit{covariate shift} can affect transfer in the high-dimensional linear regression setting. We relate each property to the bias and variance of multi-task learning and explain three negative effects with decreased task similarity,	increased sample ratio, and covariate shift under increased sample ratio. We validate the three effects on text classification tasks. Inspired by our theory, we show two practical connections of interest.
	First, single-task results can help understand when multi-task learning gives gains. Second, incrementally adding training data can mitigate negative transfer and improve multi-task training efficiency.
\end{abstract}

@@

%\textbf{Setup and Main Results.}
In this work, we study the bias and variance of multi-task learning in the high-dimensional linear regression setting \cite{HMRT19,BLLT20}.
Our key observation is that three properties of task data, including \textit{task similarity}, \textit{sample ratio}, and \textit{covariate shift}, can affect whether multi-task learning outperforms single-task learning (which we refer to as \textit{positive transfer}).
As an example, we vary each property in Figure \ref{fig_model_shift_phasetrans} for two linear regression tasks and measure the improvement of multi-task learning over single-task learning for a particular task.
We observe that the effect of transfer can be positive or negative as we vary each property.
These phenomena cannot be explained using previous techniques \cite{WZR20}.
The high-dimensional linear regression setting allows us to measure the three properties precisely.
We define each property for the case of two tasks and our definition applies to general settings.
We refer to the first task as the source task and the second as the target task.
\squishlist
	\item \textbf{Task similarity:} Assume that both tasks follow a linear model with parameters $\beta_1, \beta_2\in\real^p$, respectively.
	We measure the distance between them by $\norm{\beta_1 - \beta_2}$.
	\item \textbf{Sample ratio:} Let $n_1 = \rho_1 \cdot p, n_2 = \rho_2 \cdot p$ be the sample size of each task, where $\rho_1, \rho_2>1$ are both fixed values that do not grow with $p$.
	We measure the source/target sample ratio by $\rho_1 / \rho_2$.
%	Importantly, $\rho_2$ can be a small constant (say $2$) to capture the need for more labeled data.
	\item \textbf{Covariate shift:} Assume that the task features are random vectors with positive semidefinite covariance matrix $\Sigma_1,\Sigma_2\in\real^{p\times p}$, respectively.
	%$x = \Sigma_i^{1/2}z$, where $z\in\real^p$ consists of i.i.d. entries with mean zero and unit variance, and is a positive semidefinite matrix.
	We measure covariate shift with matrix $\Sigma_1^{1/2}\Sigma_2^{-1/2}$.
\squishend

### 2020-08-20 arxiv/proof_examples.tex
wip

\section{Proofs for Isotropic and Covariate Shifted Settings}\label{app_iso_cov}

@@

@@

\subsection{Sample Size} \label{app_proof_32}
We first prove Proposition \ref{prop_data_size}, which describes the effect of source/task data ratio on the information transfer.

>>>>

\section{Proof of the Special Case}\label{app_iso_cov}

@@

\subsection{Sample Ratio} \label{app_proof_32}
We first prove Proposition \ref{prop_data_size}, which describes the effect of source/task data ratio on the information transfer.

\subsection{Denoising}

@@

\subsection{Data Efficiency}
